---
title: "WQP Data Download"
output: html_notebook
---
# Set-Up R Environment
```{r setup}
## Libraries
library(dataRetrieval)
library(tidyverse)
library(openxlsx)
library(leaflet)                                           
library(reshape2)
library(knitr)

#Set working directory
opts_knit$set(root.dir = "C:/Users/jbehr/OneDrive - Duke University/Documents/5. R/2._WQP-EllerbeC-FlatR")
```

# Download Data
Note that the data was previously downloaded from the WQP, however here a file previously saved is used (code to download directly from the WQP is commented out).

*This will take a few minutes*
```{r}
## Clear environment
#rm(list = ls())

#Upper_Neuse_Raw_Data <- readWQPdata(huc = "03020201") # Takes awhile to load (downloaded on 7/1/2020)
#write.csv(Upper_Neuse_Raw_Data, "output/Data/WQP_Upper-Neuse_RAW.csv")

Upper_Neuse_Raw_Data<-read.csv("output/Data/WQP_Upper-Neuse_RAW.csv") 

Upper_Neuse_Water_Raw<-Upper_Neuse_Raw_Data %>% 
  filter(ActivityMediaName == "Water")
```

# Addition of Data Qualifiers
Additional columns of information is added, providing information about the sites and parameters.
```{r}
# Add long/lat for sites
Upper_Neuse_Sites<-whatWQPsites(huc = "03020201") %>% 
  select(MonitoringLocationIdentifier, MonitoringLocationName, 
         MonitoringLocationTypeName, LatitudeMeasure, 
         LongitudeMeasure, StateCode, CountyCode, HUCEightDigitCode)

Upper_Neuse_Water<-left_join(Upper_Neuse_Water_Raw, Upper_Neuse_Sites, 
                              by = "MonitoringLocationIdentifier")

# Add parameter code information
Parameter_Data <- parameterCdFile %>% 
  dplyr::rename(USGSPCode = parameter_cd)

Upper_Neuse_Water$USGSPCode<- as.character(Upper_Neuse_Water$USGSPCode)
Upper_Neuse_Water$USGSPCode<-str_pad(Upper_Neuse_Water$USGSPCode, 5, pad="0") #Makes analyses easier later

Upper_Neuse_Water<-left_join(Upper_Neuse_Water, Parameter_Data, by="USGSPCode")

# Add decade & year
Upper_Neuse_Water<-Upper_Neuse_Water %>% 
  mutate(Year = as.numeric(format(as.Date(ActivityStartDate, 
                                          format = "%Y-%m-%d"), "%Y"))) %>% 
  mutate(Decade = Year - Year %% 10)
```

# Sub-sect Data
Data is transformed into a few smaller dataframes for easier analysis.
```{r}
# Subsect data into Flat and Ellerbe River
Flat_River_Water<-Upper_Neuse_Water %>% 
  filter(str_detect(MonitoringLocationName, "FLAT")) %>% 
  mutate(N = "1") %>% 
  mutate(Site_Group = "Flat River")

Ellerbe_Creek_Water<-Upper_Neuse_Water %>% 
  filter(str_detect(MonitoringLocationName, "ELLERBE")) %>% 
  mutate(N = "1") %>% 
  mutate(Site_Group = "Ellerbe Creek")

# Merge into combined set
Both_Water<-full_join(Flat_River_Water, Ellerbe_Creek_Water)

Both_Water$Unique_YearSiteParameter_Code<-paste(Both_Water$Year, "_", 
                                                Both_Water$MonitoringLocationIdentifier, "_", 
                                                Both_Water$USGSPCode)
Both_Water$Parameter_with_units<-paste(Both_Water$USGSPCode, "_", 
                                                Both_Water$CharacteristicName, "(", 
                                                Both_Water$parameter_units, ")")

Ellerbe_Flat_Sites<-Both_Water %>% 
  group_by(MonitoringLocationIdentifier) %>% 
  tally() %>% 
  arrange(desc(n)) %>% 
  left_join(Upper_Neuse_Sites)
```

# Save Data Locally
```{r}
# Export the data or do some simple tests
write.csv(Upper_Neuse_Water, "output/Data/WQP_Upper-Neuse_WATER.csv")

write.csv(Flat_River_Water, "output/Data/WQP_FR_WATER.csv")
write.csv(Ellerbe_Creek_Water, "output/Data/WQP_EC_WATER.csv")

write.csv(Upper_Neuse_Sites, "output/Data/WQP_Upper-Neuse_SITES.csv")
write.csv(Ellerbe_Flat_Sites, "output/Data/WQP_FR-EC_SITES.csv")

#tally(group_by(Upper_Neuse_Raw_Data_Raw, ActivityMediaName)) # Just to validate
#tally(group_by(Parameter_Data, parameter_group_nm)) # Count of parameters, per group
```


