---
title: "WQP Data Download"
output: html_notebook
---
# Set-Up R Environment
```{r setup}
## Clear environment
rm(list = ls())

## Libraries
library(dataRetrieval)
library(tidyverse)
library(openxlsx)
library(leaflet)                                           
library(reshape2)
library(knitr)
library(MazamaSpatialUtils)

#Set working directory
opts_knit$set(root.dir = "C:/Users/jbehr/OneDrive - Duke University/Documents/5. R/2._WQP-EllerbeC-FlatR")
```

# Water Data

### Download Data
Note that the data was previously downloaded from the WQP, however here a file previously saved is used (code to download directly from the WQP is commented out).

*This will take a few minutes*
```{r}
#Upper_Neuse_Raw_Data <- readWQPdata(huc = "03020201") # Takes awhile to load (downloaded on 7/1/2020)
#write.csv(Upper_Neuse_Raw_Data, "output/Data/WQP_Upper-Neuse_RAW.csv")

Upper_Neuse_Raw_Data<-read.csv("output/Data/WQP_Upper-Neuse_RAW.csv") 

Upper_Neuse_Water_Raw<-Upper_Neuse_Raw_Data %>% 
  filter(ActivityMediaName == "Water")
```

### Addition of Data Qualifiers
Additional columns of information is added, providing information about the sites and parameters.
```{r}
# Add long/lat for sites
Upper_Neuse_Sites<-whatWQPsites(huc = "03020201") %>% 
  select(MonitoringLocationIdentifier, MonitoringLocationName, 
         MonitoringLocationTypeName, LatitudeMeasure, 
         LongitudeMeasure, StateCode, CountyCode, HUCEightDigitCode)

Upper_Neuse_Water<-left_join(Upper_Neuse_Water_Raw, Upper_Neuse_Sites, 
                              by = "MonitoringLocationIdentifier")

# Add parameter code information
Parameter_Data <- parameterCdFile %>% 
  dplyr::rename(USGSPCode = parameter_cd)

Upper_Neuse_Water$USGSPCode<- as.character(Upper_Neuse_Water$USGSPCode)
Upper_Neuse_Water$USGSPCode<-str_pad(Upper_Neuse_Water$USGSPCode, 5, pad="0") #Makes analyses easier later

Upper_Neuse_Water<-left_join(Upper_Neuse_Water, Parameter_Data, by="USGSPCode")

# Add decade & year
Upper_Neuse_Water<-Upper_Neuse_Water %>% 
  mutate(Year = as.numeric(format(as.Date(ActivityStartDate, 
                                          format = "%Y-%m-%d"), "%Y"))) %>% 
  mutate(Decade = Year - Year %% 10)
```

### Sub-sect Data
Data is transformed into a few smaller dataframes for easier analysis.
```{r}
# Subsect data into Flat and Ellerbe River
Flat_River_Water<-Upper_Neuse_Water %>% 
  filter(str_detect(MonitoringLocationName, "FLAT")) %>% 
  mutate(N = "1") %>% 
  mutate(Site_Group = "Flat River")

Ellerbe_Creek_Water<-Upper_Neuse_Water %>% 
  filter(str_detect(MonitoringLocationName, "ELLERBE")) %>% 
  mutate(N = "1") %>% 
  mutate(Site_Group = "Ellerbe Creek")

# Merge into combined set
Both_Water<-full_join(Flat_River_Water, Ellerbe_Creek_Water)

Both_Water$Unique_YearSiteParameter_Code<-paste(Both_Water$Year, "_", 
                                                Both_Water$MonitoringLocationIdentifier, "_", 
                                                Both_Water$USGSPCode)
Both_Water$Parameter_with_units<-paste(Both_Water$USGSPCode, "_", 
                                                Both_Water$CharacteristicName, "(", 
                                                Both_Water$parameter_units, ")")

Ellerbe_Flat_Sites<-Both_Water %>% 
  group_by(MonitoringLocationIdentifier) %>% 
  tally() %>% 
  arrange(desc(n)) %>% 
  left_join(Upper_Neuse_Sites)

# Add columns to data for plots (for 2014 data)
Both_Water_Plot_Data<-Both_Water %>% #Add average measure
  group_by(Year, MonitoringLocationIdentifier, USGSPCode) %>% 
  mutate(Average_Measure_YearLocationAnalyte=mean(ResultMeasureValue, na.rm=TRUE)) %>% 
  ungroup() %>% 
  select(Year, MonitoringLocationIdentifier, MonitoringLocationName, USGSPCode, Parameter_with_units, CharacteristicName, ActivityStartDate, ActivityStartTime.Time, Average_Measure_YearLocationAnalyte, ResultMeasureValue, parameter_units)

Both_Water_Plot_Data_2014<-Both_Water_Plot_Data %>% #Add daily flow (m3/sec)
  filter(USGSPCode == 30209) %>% #right_join below brings all data back
  rename(Stream_Flow_Value_m3persec=ResultMeasureValue) %>% 
  select(Year, MonitoringLocationIdentifier, MonitoringLocationName, ActivityStartDate, Stream_Flow_Value_m3persec) %>% 
  right_join(Both_Water_Plot_Data, 
             by= c("Year", "MonitoringLocationIdentifier", "MonitoringLocationName", "ActivityStartDate")) %>% #Brings back all the data
  filter(Year == 2014)

Param_Groups<-Parameter_Data %>%#Add parameter information
  select(USGSPCode, parameter_group_nm)
Both_Water_Plot_Data_2014<-left_join(Both_Water_Plot_Data_2014, Param_Groups)

Both_Water_Plot_Data_2014<-Both_Water_Plot_Data_2014 %>%#Add max value
  group_by(MonitoringLocationIdentifier, Parameter_with_units) %>% 
  mutate(Max_Param_Value = max(ResultMeasureValue)) %>% 
  ungroup()

Two_Sites_2014<-Both_Water_Plot_Data_2014 %>% #Select 2 sites for 2 years
  filter(str_detect(MonitoringLocationIdentifier, 
                    c("02085500|02086849"))) %>% 
  filter(Year == 2014)

Both_Water_Plot_Data_2014<-Both_Water_Plot_Data_2014 %>% #Convert to common g/L measure
  mutate(Measured_Value_gperL = case_when(str_detect(parameter_units, "mg/l") ~ Max_Param_Value/1000,
                                          parameter_units == "ug/l" ~ Max_Param_Value/1000000,
                                          parameter_units == "ng/l" ~ Max_Param_Value/1000000000)) %>%
  mutate(Measured_Value_ugperL = Measured_Value_gperL * 1000000) %>% 
  mutate(Measured_Value_ugperL_not_min = case_when(str_detect(parameter_units, "mg/l") ~ ResultMeasureValue/1000,
                                                  parameter_units == "ug/l" ~ ResultMeasureValue/1000000,
                                                  parameter_units == "ng/l" ~ ResultMeasureValue/1000000000)) %>% 
  mutate(Detect_Status = ifelse(is.na(ResultMeasureValue), "Non-Detect", "Detect"))

```

### STORET Site Data
```{r}
Ellerbe_Gorman_Data<-readWQPdata(siteid= c("21NC02WQ-J1330000", "21NC03WQ-J1330000")) %>% # EPA site near Gorman
  filter(ActivityMediaName == "Water")

Flat_Dam_Data<-readWQPdata(siteid= c("21NC02WQ-J1100000", "21NC03WQ-J1100000")) %>% # EPA site near Dam
  filter(ActivityMediaName == "Water")

STORET_Data_EC_FR<-full_join(Ellerbe_Gorman_Data, Flat_Dam_Data)
```


### Save Data Locally
```{r}
# Export the data or do some simple tests
write.csv(Upper_Neuse_Water, "output/Data/WQP_Upper-Neuse_WATER.csv")
write.csv(Both_Water, "output/Data/WQP_FR-EC_WATER.csv")

write.csv(Flat_River_Water, "output/Data/WQP_FR_WATER.csv")
```


```{r}
write.csv(Ellerbe_Creek_Water, "output/Data/WQP_EC_WATER.csv")

write.csv(Upper_Neuse_Sites, "output/Data/WQP_Upper-Neuse_SITES.csv")
write.csv(Ellerbe_Flat_Sites, "output/Data/WQP_FR-EC_SITES.csv")

write.csv(Both_Water_Plot_Data_2014, "output/Data/WQP_FR-EC_WATER-2014.csv")

write.csv(STORET_Data_EC_FR, "output/Data/WQP_FR-EC_WATER_STORET.csv")
#tally(group_by(Upper_Neuse_Raw_Data_Raw, ActivityMediaName)) # Just to validate
#tally(group_by(Parameter_Data, parameter_group_nm)) # Count of parameters, per group
```

# Biological data.
Note that data is from USGS' Biodata website. The data was downloaded manually and is brought in from a local copy.
```{r}
setSpatialDataDir("C:/Users/jbehr/OneDrive - Duke University/Documents/5. R/3. Mapping/data_raw/Data_HUCs")
loadSpatialData("WBDHU8.RData") # Download from http://data.mazamascience.com/MazamaSpatialUtils/Spatial/

setwd("C:/Users/jbehr/OneDrive - Duke University/Documents/5. R")

Inverts<-read.csv("3. Mapping/data_raw/Data_Biodata_NC/20200420.1152.InvertResults.csv")
Sites<-read.csv("3. Mapping/data_raw/Data_Biodata_NC/20200420.1152.SiteInfo.csv")

Inverts_Upper_Neuse_raw<-left_join(Inverts, Sites, by = "SiteNumber")

Inverts_Upper_Neuse_raw<- Inverts_Upper_Neuse_raw%>% 
  mutate(HUC8_FUll_Name = getHUCName(as.numeric(Inverts_Upper_Neuse_raw$Longitude_dd), as.numeric(Inverts_Upper_Neuse_raw$Latitude_dd), dataset = "WBDHU8")) %>% 
  mutate(HUC8_Number = getHUC(as.numeric(Inverts_Upper_Neuse_raw$Longitude_dd), as.numeric(Inverts_Upper_Neuse_raw$Latitude_dd), dataset = "WBDHU8"))%>% 
  filter(HUC8_Number == "03020201")

colnames(Inverts_Upper_Neuse_raw)[1]<-"SIDNO" # Need to rename, some glitch with how the first column's name comes in

Inverts_Upper_Neuse<- Inverts_Upper_Neuse_raw %>% 
  filter(FieldComponent == "M") %>% 
  rename(Field_Site = SiteName.x)

#rm(Inverts_Upper_Neuse_raw, Inverts, Sites, WBDHU8)
```

### Save Bio data

```{r}
write.csv(Inverts_Upper_Neuse, "output/Data/Upper-Neuse_Inverts.csv")
```

# Monthly Flow Data

```{r}
Flat<-read.table("data_raw/Flat_Dam.txt", header = TRUE, sep = "\t") %>% 
  mutate(Gage_ft = ifelse(parameter_cd=="65", mean_va, NA)) %>% 
  mutate(Discharge_cu3sec = ifelse(parameter_cd=="60", mean_va, NA)) %>% 
  unite(year_month, c("year_nu","month_nu")) %>% 
  select(-c(mean_va, ts_id))

Flat_1<-filter(Flat, parameter_cd == "60") %>% select(-Gage_ft)
Flat_2<-filter(Flat, parameter_cd == "65") %>% select(-Discharge_cu3sec)
Flat<-full_join(Flat_1, Flat_2, by = c("agency_cd", "site_no", "year_month")) %>% 
  select(-contains("parameter")) 

Club<-read.table("data_raw/Ellerbe_Club.txt", header = TRUE, sep = "\t") %>% 
  mutate(Gage_ft = ifelse(parameter_cd=="65", mean_va, NA)) %>% 
  mutate(Discharge_cu3sec = ifelse(parameter_cd=="60", mean_va, NA)) %>% 
  unite(year_month, c("year_nu","month_nu")) %>% 
  select(-c(mean_va, ts_id))

Club_1<-filter(Club, parameter_cd == "60") %>% select(-Gage_ft)
Club_2<-filter(Club, parameter_cd == "65") %>% select(-Discharge_cu3sec)
Club<-full_join(Club_1, Club_2, by = c("agency_cd", "site_no", "year_month")) %>% 
  select(-contains("parameter"))

Glennstone<-read.table("data_raw/Ellerbe_Glennstone.txt", header = TRUE, sep = "\t") %>% 
  mutate(Gage_ft = ifelse(parameter_cd=="65", mean_va, NA)) %>% 
  mutate(Discharge_cu3sec = ifelse(parameter_cd=="60", mean_va, NA)) %>% 
  unite(year_month, c("year_nu","month_nu")) %>% 
  select(-c(mean_va, ts_id))

Glennstone_1<-filter(Glennstone, parameter_cd == "60") %>% select(-Gage_ft)
Glennstone_2<-filter(Glennstone, parameter_cd == "65") %>% select(-Discharge_cu3sec)
Glennstone<-full_join(Glennstone_1, Glennstone_2, by = c("agency_cd", "site_no", "year_month")) %>% 
  select(-contains("parameter"))

Haw<-read.table("data_raw/Haw_Near-Jordan-Lake.txt", header = TRUE, sep = "\t") %>% 
  mutate(Gage_ft = ifelse(parameter_cd=="65", mean_va, NA)) %>% 
  mutate(Discharge_cu3sec = ifelse(parameter_cd=="60", mean_va, NA)) %>% 
  unite(year_month, c("year_nu","month_nu")) %>% 
  select(-c(mean_va, ts_id))

Haw_1<-filter(Haw, parameter_cd == "60") %>% select(-Gage_ft)
Haw_2<-filter(Haw, parameter_cd == "65") %>% select(-Discharge_cu3sec)
Haw<-full_join(Haw_1, Haw_2, by = c("agency_cd", "site_no", "year_month")) %>% 
  select(-contains("parameter")) 

All_sites_flow<-full_join(Flat, Club)
All_sites_flow<-full_join(All_sites_flow, Glennstone) 
All_sites_flow<-full_join(All_sites_flow, Haw) %>% 
  mutate(site_no=recode(site_no,
                        "2086500" = "Flat River Near Dam",
                        "208675010" = "Ellerbe Creek by Club Blvd",
                        "2086849" = "Ellerbe Creek by Glennstone",
                        "2096960" = "Haw River upstream/near Jordan Lake"))
```

### Save Flow data

```{r}
write.csv(All_sites_flow, "output/Data/Upper-Neuse_Inverts.csv")
```



